Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-13 papers/FX/draft.tex
I think we have said all that there is to be said in the case of t-distributed X.

\subsubsection{log-utility function}
We first consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:power_utility} as $\gamma \to 0$.
\begin{theorem}
  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
  function $u(\cdot) = \ln(\cdot)$, then
  \[
  \pd{G(F)}{\nu} < 0
  \]
  where $F(x, \nu)$ is the Student's t-distribution function.
\end{theorem}
\begin{proof}
  By the same argument as in the proof of theorem \ref{thrm:I}, it
  suffices to show
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &<& 0
  \end{eqnarray*}
  where $\hat\phi$ is defined in \eqref{eq:xxie2}. We have
  \begin{eqnarray*}
    && \mathcal U(F, \phi) \\
    &=&
    \int_{-\infty}^\infty
    u[(1 - \phi) e^r + \phi e^x] (1 + b \1{x < q}) f(x, \nu) dx
  \end{eqnarray*}

\end{proof}

Then it follows
\begin{eqnarray}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  } \nonumber \\
  &:=& {
    P(x^2, \nu)
    \over
    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  }
  \label{eq:xxie4.1}
\end{eqnarray}

While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
always positive, its nummerator $P(x^2, \nu)$ has a single root:
\begin{equation}
  \label{eq:xxie5}
  x_0^2 = \nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu
\end{equation}
where $W(\cdot)$ is the principle branch of the Lambert $W$
function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
the right side of \eqref{eq:xxie5} for positivity, we first note
$c'(\nu) > 0$:
\[
c'(\nu) = {
  \pi \Gamma(\nu/2 + 1/2) \left\{
    \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
  \right\}
  \over
  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
}
\]
where $\Psi(\cdot)$ is the digamma function:
\[
\Psi(x) = \td{\ln[\Gamma(x)]}{x}
\]
\begin{minipage}{0.48\textwidth}
  $\Psi(x)$ \linebreak
  \includegraphics[width=\textwidth]{digamma.png}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
  As shown in the figure to the left, $\Psi(x)$ is increasing
  for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
  So we have
  \begin{eqnarray*}
    && \nu \left[
    \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
    \right] - 1 \\
    &\geq& 1 \times \left[
    \Psi(1/2 + 1/2) - \Psi(1/2)
    \right] - 1 \\
    &=& \ln(4) - \ln(e) \\
    &>& 0
  \end{eqnarray*}
  Thus $c'(\nu) > 0$. Furthermore, we recall
  $W(\cdot)$ is increasing on its principle branch. So
\end{minipage}
\begin{eqnarray*}
  &&
  W\left[
    -\left( 1 + {1 \over \nu} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &>& 
  W\left[
    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &=& W(-y e^{-y}) + y
\end{eqnarray*}
where
\[
y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
\]
Now notice
\[
\ln(y e^{-y}) = \ln(y) - y
\]
is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
increasing function. Hence we have
\begin{eqnarray*}
  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
\end{eqnarray*}
Now it is clear
\[
\nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu > 0
\]
Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
single positive root, it remains to determine the sign of
$\pd{f}{\nu}(x, \nu)$ on the two sides of the root in order to apply
lemma \ref{lemma:I}. For this purpose we observe
\begin{equation}
  \label{eq:xxie5.1}
  P(0, \nu) = 2 \nu^2 c'(\nu) > 0
\end{equation}
So we want to investigate $\pd{P}{x^2}(x^2, \nu)$:
\begin{eqnarray}
  && \pd{P}{x^2}(x^2, \nu) \nonumber \\
  &=&
  2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
  1 + {x^2 \over \nu}
  \right)
  \label{eq:xxie6}
\end{eqnarray}
From \eqref{eq:xxie5.1} and \eqref{eq:xxie6} it is clear
$P(x^2, \nu) > 0$ for $0 \leq x^2 < x_0^2$ and
$P(x^2, \nu) < 0$ for $x^2 > x_0^2$.

\section{Review of multivariate GARCH models:
  same tail index or not}
Many multivariate GARCH models imply a common tail index for all
series under consideration, e.g. the Orthogonal GARCH model of
Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
well as the Full Factor GARCH model of Vrontos et al (2003)
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process. As a result of this treatment, each and every series under
consideration necessarily has the lowest of the tail indices of its
contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
as Andersen, Davis, Kreiss and Mikosch (2009)
\cite{andersen2009handbook}).

@@

\section{Proof of lemma \ref{lemma:II}}

>>>>

\subsubsection{log-utility function}
We first consider the case where the utility function is simply
$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
\eqref{eq:power_utility} as $\gamma \to 0$. Straightforward
computation gives
\begin{eqnarray*}
  && \mathcal U(F, \phi) \\
  &=&
  \int_{0}^{\infty}
  \left[
    \ln(C(x)) + (1 + b)\ln(C(-x))
  \right]
  f(x, \nu) dx
  + b \int_{0}^{q}
  \ln(C(x))
  f(x, \nu) dx \\
  &&
  - b \ln(C(q)) F(q, \nu)
\end{eqnarray*}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}. Notice that
when $b = 0$, $\mathcal U(F, \phi)$  reduces to $\E u(C(X))$.
Also observe $\ln(C(x)C(-x))$ is an increasing function:
\begin{eqnarray*}
  \pd{C(x) C(-x)}{x}
  &=&
  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
\end{eqnarray*}
Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
investors guided by this utility function would be seeking the
smallest $\nu$, i.e. the heaviest tail in the market.

However, this is not necessarily the case when $b > 0$. For example,
consider the case when $q = 0$ and $\hat\phi = 1$. This may be
viewed an approximation to the situation where the interest rate $r$ is
vanishingly low while the investor is more conservative than
greedy. In this case we have
\begin{eqnarray*}
  \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
\end{eqnarray*}
Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
$\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
would rather seek the largest $\nu$ or in other words, the lightest
tail in the market.

%   where
%   \[
%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
%   \]
%   Now notice
%   \[
%   \ln(y e^{-y}) = \ln(y) - y
%   \]
%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
%   function. Hence we have
%   \begin{eqnarray*}
%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
%   \end{eqnarray*}
%   Now it is clear
%   \[
%   \nu\exp\left\{
%   W\left[
%     -\left(1 + {1 \over \nu}\right)
%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
%   \right]
%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
%   &=& W(-y e^{-y}) + y
% \end{eqnarray*}
\section{Review of multivariate GARCH models:
  same tail index or not}
Many multivariate GARCH models imply a common tail index for all
series under consideration, e.g. the Orthogonal GARCH model of
Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
well as the Full Factor GARCH model of Vrontos et al (2003)
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process. As a result of this treatment, each and every series under
consideration necessarily has the lowest of the tail indices of its
contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
as Andersen, Davis, Kreiss and Mikosch (2009)
\cite{andersen2009handbook}).

@@

\section{Proof of lemma \ref{lemma:II}}
\begin{proof}
  Straightforward computation gives
  \begin{eqnarray}
    \pd{f(x, \nu)}{\nu} &=& {
      c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
      -
      \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
      \over
      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
    } \nonumber \\
    &:=& {
      P(x^2, \nu)
      \over
      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
    }
    \label{eq:xxie4.1}
  \end{eqnarray}

  While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
  always positive, its nummerator $P(x^2, \nu)$ has a single root:
  \begin{equation}
    \label{eq:xxie5}
    x_0^2 = \nu\exp\left\{
      W\left[
        -\left(1 + {1 \over \nu}\right)
        e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
      \right]
      + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
    \right\} - \nu
  \end{equation}
  where $W(\cdot)$ is the principle branch of the Lambert $W$
  function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
  the right side of \eqref{eq:xxie5} for positivity, we first note
  $c'(\nu) > 0$:
  \[
  c'(\nu) = {
    \pi \Gamma(\nu/2 + 1/2) \left\{
      \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
    \right\}
    \over
    2 \Gamma(\nu/2) (\pi \nu)^{3/2}
  }
  \]
  where $\Psi(\cdot)$ is the digamma function:
  \[
  \Psi(x) = \td{\ln[\Gamma(x)]}{x}
  \]
  \begin{minipage}{0.48\textwidth}
    $\Psi(x)$ \linebreak
    \includegraphics[width=\textwidth]{digamma.png}
  \end{minipage}\hfill
  \begin{minipage}{0.5\textwidth}
    As shown in the figure to the left, $\Psi(x)$ is increasing
    for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
    So we have
    \begin{eqnarray*}
      && \nu \left[
        \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
      \right] - 1 \\
      &\geq& 1 \times \left[
        \Psi(1/2 + 1/2) - \Psi(1/2)
      \right] - 1 \\
      &=& \ln(4) - \ln(e) \\
      &>& 0
    \end{eqnarray*}
    Thus $c'(\nu) > 0$. Furthermore, we recall
    $W(\cdot)$ is increasing on its principle branch. So
  \end{minipage}
  \begin{eqnarray*}
    &&
    W\left[
      -\left( 1 + {1 \over \nu} \right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
    &>& 
    W\left[
      -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
    &=& W(-y e^{-y}) + y
  \end{eqnarray*}
  where
  \[
  y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
  \]
  Now notice
  \[
  \ln(y e^{-y}) = \ln(y) - y
  \]
  is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
  increasing function. Hence we have
  \begin{eqnarray*}
    W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
  \end{eqnarray*}
  Now it is clear
  \[
  \nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu > 0
  \]
  Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
  single positive root, it remains to determine the sign of
  $\pd{f}{\nu}(x, \nu)$ on the two sides of the root. For this purpose
  we observe
  \begin{equation}
    \label{eq:xxie5.1}
    P(0, \nu) = 2 \nu^2 c'(\nu) > 0
  \end{equation}
  So we want to investigate $\pd{P}{x}(x, \nu)$:
  \begin{eqnarray}
    \pd{P}{x}(x, \nu) &=&
    2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
      1 + {x \over \nu}
    \right)
    \label{eq:xxie6}
  \end{eqnarray}
  Clearly, $\pd{P}{x}(0, \nu) > 0$. Hence from \eqref{eq:xxie5.1} and
  \eqref{eq:xxie6} it is clear
  \begin{equation*}
    \text{sign}\left[
      \pd{f}{\nu}(x, \nu)
    \right]
    = \left\{
    \begin{array}{rl}
      1 & 0 < x < x_0 \\
      -1 & x > x_0
    \end{array}
    \right.
  \end{equation*}
  where $x_0$ is the positive root of \eqref{eq:xxie5}.
\end{proof}
